### 2017-01-26 paper/benchmarks.tex
wip

\section{Benchmarks}
To check performance of our new parallel arrow API, we conducted some benchmarks. These are based on a sudoku solver from the examples for Simon Marlows book "Parallel and Concurrent Programming in Haskell" which can be found on his GitHub \cite{marlow_parconc}. The results are displayed in the following graphs which are ordered by problem size.
\\\\
The Benchmarks were run on a Core i7-3970X CPU @ 3.5GHz with 6 cores and 12 threads. All versions were run 100 times. The resulting running-times were then averaged.
\\
\begin{tikzpicture}
\begin{axis}[
xbar, xmin=0,
width=12cm, height=7cm, enlarge y limits=0.5,
xlabel={avg. seconds run on sudoku17.1000.txt (100 runs)},
symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
ytick=data,
nodes near coords, nodes near coords align={horizontal},
]
\addplot coordinates {(.215,parrows-sudoku-farmChunk) (.213,parrows-sudoku-farm) (.188,parrows-sudoku-parmapstream) (.197,parrows-sudoku-parmap) (.527,parmonad-sudoku-parmap) (1.338,sudoku-sequential)};
\end{axis}
\end{tikzpicture}
\textit{Note: the bad result for parmonad-sudoku-parmap seems to be a artefact as it performs much better in the other benchmarks}
\\\\
\begin{tikzpicture}
\begin{axis}[
xbar, xmin=0,
width=12cm, height=7cm, enlarge y limits=0.5,
xlabel={avg. seconds run on sudoku17.16000.txt (100 runs)},
symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
ytick=data,
nodes near coords, nodes near coords align={horizontal},
]
\addplot coordinates {(3.218,parrows-sudoku-farmChunk) (3.148,parrows-sudoku-farm) (2.755,parrows-sudoku-parmapstream) (2.860,parrows-sudoku-parmap) (2.936,parmonad-sudoku-parmap) (21.324,sudoku-sequential)};
\end{axis}
\end{tikzpicture}
\begin{tikzpicture}
\begin{axis}[
xbar, xmin=0,
width=12cm, height=7cm, enlarge y limits=0.5,
xlabel={avg. seconds run on sudoku17.49151.txt (100 runs)},
symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
ytick=data,
nodes near coords, nodes near coords align={horizontal},
]
\addplot coordinates {(10.038,parrows-sudoku-farmChunk) (10.151,parrows-sudoku-farm) (8.776,parrows-sudoku-parmapstream) (8.969,parrows-sudoku-parmap) (8.679,parmonad-sudoku-parmap) (60.931,sudoku-sequential)};
\end{axis}
\end{tikzpicture}
\\\\
As we can see, the parallel arrow versions of the program all have around the same speedup as the ParMonad based version. This means that using the \code{ArrowParallel} typeclass doesn't add any real overhead. Also, the slightly worse results for "parrows-sudoku-farm" and "parrows-sudoku-farmChunk" can be explained by either imperfect parameters or by the fact that the benchmarks do not really require the scheduling introduced by the skeletons which in this case introduces unnecessary overhead. This would probably look different if the benchmark would use e.g. a divide and conquer approach to solve the sudoku puzzles.

>>>>

\section{Benchmarks}
To check the performance of our new parallel arrow API, we conducted some benchmarks. These are based on a sudoku solver from the examples for Simon Marlows book "Parallel and Concurrent Programming in Haskell" which can be found on his GitHub \cite{marlow_parconc}. The results are displayed in the following graphs which are ordered by problem size.
\\\\
The Benchmarks were run on a Core i7-3970X CPU @ 3.5GHz with 6 cores and 12 threads. For sake of comparability with Simon Marlow's parallel version which uses the ParMonad, we use the ParMonad backend for the parallel arrow versions as well.
\\\\
\framebox{
	\begin{tikzpicture}[thick, scale=0.7]
	\begin{axis}[
	xbar, xmin=0,
	width=12cm, height=7cm, enlarge y limits=0.5,
	xlabel={avg. seconds run on sudoku17.1000.txt (100 runs)},
	symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
	ytick=data,
	nodes near coords, nodes near coords align={horizontal},
	]
	\addplot coordinates {(.215,parrows-sudoku-farmChunk) (.213,parrows-sudoku-farm) (.188,parrows-sudoku-parmapstream) (.197,parrows-sudoku-parmap) (.527,parmonad-sudoku-parmap) (1.338,sudoku-sequential)};
	\end{axis}
	\end{tikzpicture}
}
\\\\
\textit{Note: the bad result for parmonad-sudoku-parmap seems to be a artefact as it performs much better in the other benchmarks}
\\\\
\framebox{
	\begin{tikzpicture}[thick, scale=0.7]
	\begin{axis}[
	xbar, xmin=0,
	width=12cm, height=7cm, enlarge y limits=0.5,
	xlabel={avg. seconds run on sudoku17.16000.txt (100 runs)},
	symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
	ytick=data,
	nodes near coords, nodes near coords align={horizontal},
	]
	\addplot coordinates {(3.218,parrows-sudoku-farmChunk) (3.148,parrows-sudoku-farm) (2.755,parrows-sudoku-parmapstream) (2.860,parrows-sudoku-parmap) (2.936,parmonad-sudoku-parmap) (21.324,sudoku-sequential)};
	\end{axis}
	\end{tikzpicture}
}
\\\\
\framebox{
	\begin{tikzpicture}[thick, scale=0.7]
	\begin{axis}[
	xbar, xmin=0,
	width=12cm, height=7cm, enlarge y limits=0.5,
	xlabel={avg. seconds run on sudoku17.49151.txt (100 runs)},
	symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
	ytick=data,
	nodes near coords, nodes near coords align={horizontal},
	]
	\addplot coordinates {(10.038,parrows-sudoku-farmChunk) (10.151,parrows-sudoku-farm) (8.776,parrows-sudoku-parmapstream) (8.969,parrows-sudoku-parmap) (8.679,parmonad-sudoku-parmap) (60.931,sudoku-sequential)};
	\end{axis}
	\end{tikzpicture}
}
\\\\
As we can see, the parallel arrow versions of the program all have around the same speedup as the ParMonad based version. This means that using the \code{ArrowParallel} typeclass doesn't add any real notable overhead. Also, the slightly worse results for "parrows-sudoku-farm" and "parrows-sudoku-farmChunk" can be explained by either imperfect parameters or by the fact that the benchmarks do not really require the scheduling introduced by the skeletons which in this case introduces unnecessary overhead. This would probably look different if the benchmark would use e.g. a divide and conquer approach to solve the sudoku puzzles.

### 2017-01-26 paper/parrows.tex

\subsection{Benefits of parallel Arrows}
We have seen that we can wrap parallel Haskells inside of the \code{ArrowParallel} interface, but why do we abstract parallelism this way and what does this approach do better than the other parallel Haskells?
\begin{itemize}
	\item \textbf{Arrow API benefits}:
	With the \code{ArrowParallel} typeclass we do not lose any benefits of using arrows as \code{parEvalN} is just yet another arrow combinator. The resulting arrow can be used in the same way its serial version can be used. This is a big advantage of this approach, especially compared to the monad solutions as all arrow combinators will continue to work instead of requiring special ones for each parallel monad.
	\item \textbf{Abstraction}:
	With the \code{ArrowParallel} typeclass, we abstracted all parallel implementation logic away from the business logic. This gives us the beautiful situation of being able to write our code against the interface the typeclass gives us without being bound to any parallel Haskell. So as an example, during development, we can run the code on the simple Multicore version and afterwards deploy it on a cluster using an Eden built, by just swapping out the actual \code{ArrowParallel} instance.
\end{itemize}

>>>>

\subsection{Benefits of parallel Arrows}
We have seen that we can wrap parallel Haskells inside of the \code{ArrowParallel} interface, but why do we abstract parallelism this way and what does this approach do better than the other parallel Haskells?
\begin{itemize}
	\item \textbf{Arrow API benefits}:
	With the \code{ArrowParallel} typeclass we do not lose any benefits of using arrows as \code{parEvalN} is just yet another arrow combinator. The resulting arrow can be used in the same way its serial version can be used. This is a big advantage of this approach, especially compared to the monad solutions as all arrow combinators will continue to work instead of requiring special ones for each parallel monad.
	\item \textbf{Abstraction}:
	With the \code{ArrowParallel} typeclass, we abstracted all parallel implementation logic away from the business logic. This gives us the beautiful situation of being able to write our code against the interface the typeclass gives us without being bound to any parallel Haskell. So as an example, during development, we can run the code on the simple Multicore version and afterwards deploy it on a cluster by converting it into an Eden version, by just swapping out the actual \code{ArrowParallel} instance.
\end{itemize}

### 2017-01-26 paper/utilityFunctions.tex

\section{Utility Functions}\label{utilfns}
Before we go into detail on parallel arrows, we introduce some utility combinators first, that will help us later: map and zipWith on arrows.
\\
The map combinator lifts any arrow \code{arr a b} to an arrow \code{arr [a] [b]} \cite{programming_with_arrows},
\begin{lstlisting}[frame=htrbl]
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
mapArr f =
	arr listcase >>>
	arr (const []) ||| (f *** mapArr f >>> arr (uncurry (:)))
	where
		listcase [] = Left ()
		listcase (x:xs) = Right (x,xs)
\end{lstlisting}
and zipWith lift any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}.
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>> mapArr f
\end{lstlisting}
These two combinators make use of the \code{ArrowChoice} typeclass, which allows us to use the \code{|||} combinator. It takes two arrows \code{arr a c} and \code{arr b c} and combines them into a new arrow \code{arr (Either a b) c} which pipes all \code{Left a}'s to the first arrow and all \code{Right b}'s to the second arrow.
\begin{lstlisting}[frame=htrbl]
(|||) :: ArrowChoice arr a c -> arr b c -> arr (Either a b) c
\end{lstlisting}
With the zipWithArr combinator we can also write a combinator \code{listApp}, which lifts a list of arrows \code{[arr a b]} to an arrow \code{arr [a] [b]}.
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) => [arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
This combinator also makes use of the \code{ArrowApply} typeclass which allows us to evaluate arrows with \code{app :: arr (arr a b, a) c}.

>>>>

\section{Utility Functions}\label{utilfns}
Before we go into detail on parallel arrows, we introduce some utility combinators first, that will help us later: map and zipWith on arrows.
\\
The mapArr combinator, that lifts any arrow \code{arr a b} to an arrow \code{arr [a] [b]} \cite{programming_with_arrows},
\begin{lstlisting}[frame=htrbl]
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
mapArr f =
	arr listcase >>>
	arr (const []) ||| (f *** mapArr f >>> arr (uncurry (:)))
	where
		listcase [] = Left ()
		listcase (x:xs) = Right (x,xs)
\end{lstlisting}
and zipWithArr, which lifts any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}.
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>> mapArr f
\end{lstlisting}
These two combinators make use of the \code{ArrowChoice} typeclass, which allows us to use the \code{|||} combinator. It takes two arrows \code{arr a c} and \code{arr b c} and combines them into a new arrow \code{arr (Either a b) c} which pipes all \code{Left a}'s to the first arrow and all \code{Right b}'s to the second arrow.
\begin{lstlisting}[frame=htrbl]
(|||) :: ArrowChoice arr a c -> arr b c -> arr (Either a b) c
\end{lstlisting}
With the zipWithArr combinator we can also write a combinator \code{listApp}, which lifts a list of arrows \code{[arr a b]} to an arrow \code{arr [a] [b]}.
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) => [arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
This combinator also makes use of the \code{ArrowApply} typeclass which allows us to evaluate arrows with \code{app :: arr (arr a b, a) c}.
